Does Anybody Really Know What Time It Is?\\ Automating the Extraction of Date Scalars

### 2015-10-15 icde2016/background.tex
Fei's latest comments

The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the formatting language defined by the ICU open-source project. We chose this format because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\eg month names). ICU's format syntax is typical of most syntax languages and provides a large number of date part codes, and a summary of it is reproduced in Table 2. The complete syntax is documented at the ICU web site~\cite{ICU}.  Other syntaxes can either be generated directly from the algorithms, or by translating the ICU syntax into into the desired syntax. (In fact, this latter approach is how our visualization system internally generates formats for RDBMSes that have non-ICU syntax.)

>>>>

The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the formatting language defined by the ICU open-source project~\cite{ICU}. We chose this format because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\eg month names). ICU's format syntax is typical of most syntax languages and provides a large number of date part codes, and a summary of it is reproduced in Table 2. The complete syntax is documented at the ICU web site.  Other syntaxes can either be generated directly from the algorithms, or by translating the ICU syntax into into the desired syntax. (In fact, this latter approach is how our visualization system internally generates formats for RDBMSes that have non-ICU syntax.)

### 2015-10-15 icde2016/mdl.tex
Fei's latest comments

\subsubsection{Domain Characteristics}
Date domains typically have small widths, so we found it advantageous to provide the shortest and longest match sizes for use in structure enumeration and matching.

Date domains are also often uniform in that adding more characters to a mismatch will not help. For example, a 2-digit day domain that does not match a 1-letter substring will not be able to generate a match by adding more characters.

@@

\subsection{Global Pruning}
The pruning rules used for the structure extraction reduce the search space dramatically, but they are also contextual and can only look backwards. The domains also contain a fair amount of ambiguity that requires the application of domain knowledge. We therefore found it necessary to add some post-generation global pruning rules:
\begin{itemize}
\item The set of date parts cannot contain place value gaps (\eg structures that have year and day without month are removed.)
\item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as \texttt{mhs}.)
\item The existence of time parts cannot make dates incomplete (\eg patterns like year-day-hour are removed.)
\item Two digit years require special handling. In particular, they cannot appear adjacent to a two-digit field if the structure contains punctuation. (This can come up in some small early 21st century year domains where a two-digit year can masquerade as almost any numeric field.)
\end{itemize}

@@

\subsection{Ranking}
MDL structure analysis naturally produces a ranked list of format candidates, but we have found that a number of other properties of the formats should be preferred over simple compactness:
\begin{itemize}
\item Because we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
\item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\eg month-day-year over hour-minute-second).
\item If two formats from different locales give the same results, prefer the original column locale. The sample set may have missed an example where this could be important.
\item If the format has an ambiguous date order (\eg all days are less than 12), then prefer the default date order of the locale. Again, the sample set may have missed a counterexample, so this is the best option.
\item Once these semantic preferences have been considered, we then prefer the more compact (MDL) representation.
\end{itemize}

>>>>

\subsubsection{Domain Characteristics}
Date domains typically have small widths, so we found it advantageous to provide the shortest and longest match sizes for use in structure enumeration and matching. For example, months have between $1$ and $2$ digits, so there is no need to test matches whose length is outside this range.

Date domains are also often uniform in that adding more characters to a mismatch will not help. For example, a $2$-digit day domain that does not match a $1$-letter substring will not be able to generate a match by adding more characters.

@@

\subsection{Global Pruning}
The pruning rules used for the structure extraction reduce the search space dramatically, but they are also contextual and can only look backwards. The domains also contain a fair amount of ambiguity that requires the application of domain knowledge. We therefore found it necessary to add some post-generation global pruning rules:
\begin{itemize}
\item The set of date parts cannot contain place value gaps (\eg structures that have year and day without month are removed.)
\item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as \texttt{mhs}.)
\item The existence of time parts cannot make dates incomplete (\eg patterns like year-day-hour are removed.)
\item Two digit years require special handling. In particular, they cannot appear adjacent to a two-digit field if the structure contains punctuation. (This can come up in some small early 21st century year domains where a two-digit year can masquerade as almost any numeric field \eg $08$ for $2008$.)
\end{itemize}

@@

\subsection{Ranking}
MDL structure analysis naturally produces a ranked list of format candidates, but we have found that a number of other properties of the formats should be preferred over simple compactness:
\begin{itemize}
\item Because we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
\item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\eg month-day-year over hour-minute-second).
\item If two formats from different locales give the same results, prefer the original column locale. The sample set may have missed an example where this could be important. For example, the month name for "September" is the same in both German and English, but "October" and "Oktober" are different.
\item If the format has an ambiguous date order (\eg all days are less than 12), then prefer the default date order of the locale. Again, the sample set may have missed a counterexample, so this is the best option.
\item Once these semantic preferences have been considered, we then prefer the more compact (MDL) representation.
\end{itemize}
